Refinement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-08-02 rigid_bodies.tex
Revision of the abstract

\begin{abstract}
Many deterministic and stochastic molecular dynamics methods that aim at reproducing a target temperature have equipartition of kinetic energy as a premise.
While this intends to mimic the real world, discretization errors, which are intrinsic to numerical integrators, may distort such assumption.
Fortunately, backward error analysis is a powerful tool that allows us to identify the quantity actually subject to equipartition, which is a shadow Hamiltonian that only coincides with the specified Hamiltonian when the time-step size approaches zero.
This paper deals with discretization effects in a straightforward way: with a minimal computational overhead, we obtain a refined kinetic energy and a refined Hamiltonian as suitable estimators of their
shadow counterparts and then we tune the thermostatting procedure with this refined kinetic energy.
The case of rigid water has served as an excellent test case for our proposal, because its numerical stability extends up to time steps large enough to yield pronounced discretization errors in Verlet-type integrators.
By applying our new approach, we have been able to correct the discretization effects in equilibrium properties for time-step sizes up to 5 fs.
\end{abstract}

>>>>

\begin{abstract}
Kinetic energy equipartition is a premise for many deterministic and stochastic molecular dynamics methods that aim at sampling a canonical ensemble.
While this is expected for real systems, discretization errors introduced by the numerical integration may distort such assumption.
Fortunately, backward error analysis allows us to identify the quantity that is actually subject to equipartition.
This is related to a shadow Hamiltonian, which coincides with the specified Hamiltonian only when the time-step size approaches zero.
This paper deals with discretization effects in a straightforward way.
With a small computational overhead, we obtain refined versions of the kinetic and potential energies, whose sum is a suitable estimator of the shadow Hamiltonian.
Then, we tune the thermostatting procedure by employing the refined kinetic energy instead of the conventional one.
This procedure is shown to reproduce a canonical ensemble compatible with the refined system, but canonical averages regarding the original system can be easily recovered by reweighting.
Water modeled as a rigid body is an excellent test case for our proposal, because its numerical stability extends up to time steps large enough to yield pronounced discretization errors in Verlet-type integrators.
By applying our new approach, we were able to mitigate discretization effects in equilibrium properties of liquid water for time-step sizes up to 5 fs.
\end{abstract}
